Rotation effects on the Lyman-$\alpha$ line morphology in distant galaxies

### 2013-05-30 paper/Analisis.tex
Conclusions

\begin{abstract}
Rotation is present in the gas kinematics of galaxies up to the
highest redshifts. In this paper we present for the first time
radiative transfer calculations that show the impact of rotation on
the morphology of the Lyman $\alpha$ line. To this end we construct
simplified models where a galaxy is modeled as an homogeneous sphere
composed as an homogenous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with linear
velocities at the surface in the range $0-300$ \kms. We consider
radiation sources both in the center of the rotating cloud and also
homogeneously distributed around the sphere. We find that higher
rotational velocities increase the width of each peak in the outgoing
line profile while it also increases the amount of Lyman alpha photons
escaping in the line center. This trends makes that for high
rotational velocities and large Hydrogen optical depths the double
peak of the line tends to be erased an be replaced by a single peak the
lines center. This is more pronounced for radiation sources
homogeneously distributed. Concerning the escape fraction we find that
rotation does not have any effect, provided that all the sources are
centrally emitted. However in the case of homogeneously emitted
sources we measure an increase of about a factor of $2$ in the escape
fraction for higher rotational velocity vealues.

Our work shows clearly that gas rotation has a non
negligeable impact on the shape of the Lyman $\alpha$ line.

@@

Due to the resonant nature of the lyman alpha line, gas kinematics
play an important role shaping its morphology. In the literature there
has been extensive studies of outflow/inflow configurations.

In this paper we study for the first time the impact of rotation on
the morphology of the Lyman $\alpha$ line. To isolate the effects of
rotation we focus on a simple system: the gas distribution is
spherical, with homogenous density and the gas rotates as a solid
body. We base our work on the Monte Carlo base radiative transfer code
CLARA \citep{CLARA}.

@@

The central result of this paper is summarized in
Fig.~\ref{fig:DifferentVelocities} where we show that rotation has a
considerable effect on the morphology of the emergent Lyman-alpha line
both in the case where the photons are emitted at the sphere's center
and when they are initialized with an homogeneous distribution all
over the gas volume.

The results for these outgoing spectra are integrated over the whole
sphere, meaning that all the escaping photons were taking into account
regardless of the direction of the outgoing photons. Figure
\ref{fig:DifferentObservers} shows how if one gives a weight to each outgoing
photon according to its direction when escaping the gas distribution
it is posible to detect notable differences in the spectrum for
different viewing angles.

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{7tDifSpeedsZ.png}
  \includegraphics[width=0.45\textwidth]{7tHOMDifSpeeds1.png}
\caption{Shape of the lyman alpha line for
    differents velocities. The left (right) panel shows the central
    (homogeneous) photon distribution. All photons were taken into
    account regardless of their outgoing direction of propagation.
  \label{fig:differentvelocities}}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.45\textwidth]{Observers7t.png}
\includegraphics[width=0.45\textwidth]{Observers7tHOM.png}
\caption{Same as Figure \ref{fig:differentvelocities} but this time
  the differen lines show the line morphology for different polar angles
  $\theta$ with respect to the rotation axis. In this case each photon
  has a weight dependent on their outgoing direction. All spectra correspond
  to the same $V_{\rm max}=300$\kms, Optical   Depth $\tau=10^{7}$ and
  Central Distribution without dust.
  \label{fig:differentobservers}}

@@

In the following subsections we quantify the trends observed in
Fig. \ref{fig:differentvelocities} and
Fig. \ref{fig:differentobservers} as a function of the maximum
rotation velocity $V_{\rm max}$ and the position of the observer with
respect to the rotation angle, $\mu = \cos{\theta}$. All the results
in this section will be expressed in terms of resframe wavelength.

@@

\subsection{Line width and peak maxima}
\label{sec:widthpeak}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{WidthvsVmax.png}
\caption{Width of the lyman-alpha line for all the models.   \label{fig:widthvsvelocity}}
\end{figure}

\begin{figure*}%QUITARLA Y COMENTARLA
    \includegraphics[width=0.45\textwidth]{WidthvsThetaCentral.png}
    \includegraphics[width=0.45\textwidth]{WidthvsTheta.png}
    \caption{Width of the lyman-alpha line for all the models. 
  \label{figure:widthvstheta}
}
\end{figure*}

The first quantitative conclusion of the effect of rotation in the
Lyman alpha line is that the double peaks in the line tend to broaden
until they reach a single broad emission peak. This is most evident in
the case of Lyman-alpha sources homogeneously distributed in the gas
distributions (Fig. \ref{differentvelocities} right panel).

@@

This result includes all the outgoing photons, regardless of the
position of the observer. In Figure~\ref{fig:widthvstheta} we
take into account the different positions of the observer in the
measurement of the half-width $W_{1/2}$. From this we conclude
that observers with a line of sight perpendicular to the axis of
rotation (i.e. edge-on in the case of spiral galaxy) tend to meaure
larger line widths than observers aligned with the rotation axis
(i.e. face-on). The influence of the observer position on the line
width, amounting always less than $15\%$ of a difference with respect
to the result that takes into account all the outgoing photons with
the same weight regardless of the relative observer position.

%Both for central and homogeneous distribution of sources, We found a
%dependency of the width with the rotation velocity of the gas cloud,
%this results are summarize in Figure ~\ref{figure:width}. As velocity
%increase also the equivalent width increase (for both homogeneous and
%central distribution), it means that some photons escape with fewer
%scatterings than the static case but at the same time a few photons escape with% more scatterings, this can be seen clearly in
%Figure. (Nscatt)\\

%We also found a dependency with the viewing angles, in particular as
%the angle increase the width also increase Figure
%~\ref{fig:withvstheta}, this was also found in Verhemme et al. 2012
%for a rotating disk but the pure effect of rotation can not been study
%as it is in our case. Following the convention of Verhamme et al 2012
%we define $\mu=cos(\theta)$, and we can make a polynomial fit for the
%$FWHM(\mu)$ for our models, this is resumed in Table ~\ref{table:fits}

@@

In Figure ~\ref{fig:maximavsvelocity} we present the position
of the maxima as a function of $V_{\rm max}$. For the photons emitted
in the line center we do not find any variation in the position of the
maxima in the range of explored parameter space. However in the
homogeneous case we can see how the maxima goes to $x_{m}=0$, meaning
that the double peak is converted into a single peak.

This transition to a single peak line occurs for the systems where it
becomes easiear for a bulk of the photons to escape with the lowest
number of scatterings possible. This can explain how the single peak
stage can be achieved in the homogeneous source distribution where
there is fraction of the photons inside a photosphere region with
$\tau_{H}\sim 1$ which allows them to escape within one
scatter. Increasing the rotational velocity makes it easier for the
photons in this photosphere region to escape.

In Figure ~\ref{fig:maximavsangle} we present the position of the
maxima as a function of the viewing angle $\mu$ for all the optical
depths and rotational velocities and the two different photon
distributions [FALTA LA GRAFICA].

@@

Finally we also report on the effect of the neutral Hydrogen
optical depth $\tau_{H}$ on the maxima position $x_{m}$
Figure~\ref{figure:maximavsOD}. We find that, at fixed
rotation velocity, the position of the maxima increases with optical
depth as expected from basic theoretical consideraions. We compare
our results with the expected theoretical escaling for an infinnite
slab.

@@

Untill know we have focus in how rotation affect the morphology of the lyman alpha line, now we turn to study deeply the posible causes of this effects. As recombination is the main cause of the lyman alpha, it is important to study the relationship between the number of times that a photon is absorbed and re emitted hereafter scatterings with it's shift in the wavelenght. \\

As a first approach we study the number of scatterings of all the photons at different velocites and for both distributions Figure~\ref{fig:Nscatt}. For the central distribution we found that the average number of scattering does not have a relevant change (put the exact value) with $V_{max}$, For the Homogeneous distribution we found that as $V_{max}$ increase the average number of scatterings decrease, it means that rotation make photons to escape easier from the cloud, and the reason why this is not observed in the central distribution. it is because the amount of gas that every photon has to break through is so large an so the average number of scatterings, making the effect of rotation unobservable.\\

In order to understand the effects of rotation in the morphology of the line we make histograms of $<N_{scatt}>$  in function of $X_{m}$ for both the central Fig. ~\ref{fig:NscattHisto} and the  homogeneous distribution Fig.~\ref{fig:NscattHistoHOM}. For the central distribution and for lower velocities we found that for some value (put the value) of $N_{scatt}$ the mayority of photons have a defined value of $X_{m}$. While for higher velocities photons tend spread out in $X_{m}$(Make position analysis of this photons).\\

For the homogenepus distribution we found the same effect but also, for highest velocities the mayority of the photons near the surface escape with much less scatterings that in the static case. This analysis let us conclude that rotation make photons escape with less scatterings and it spread out the wavelenght of the of the outgoing photons represented by $X_{m}$.

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NscattvsVmax.png}
\caption{Nscatt\label{fig:Nscatt}} 
\end{figure}

@@

We have seen in the previous section that rotation affects the number
of scatterings that ..\\

Of particular interest is to compute the escape fraction of Lyman $\alpha$ photons coming from the most distant galaxies, due to the fact that with the observed intensity of the lyman alpha line quantities as the LF and SFR can be derived (put some references). \\

Previous studies have shown the correlation of the Escape fraction with galactic mass (Lauren 2009 et al, Dayal et al 2010) abundacies and the kinematics of dust. In order to study pure rotational effects in the escape fraction we fixed the dust abundance $\tau_{A}=0.01$ and the galaxy mass. We compute the escape fraction for the models described in Table \ref{table:models}.\\

For a realistic model we also take into account the viewing angle, first we fixed the viewing angle at $\theta = 0$  Fig. and then we fixed the velocity in order to see the escape fraction correlation with the viewing angle. Therefore we define the escape fraction as:\\

@@

Where NI is the initial number of photons and NF is the final,
$\vec{k}$ is the rotation axis direction and $\vec{o}$ the observer
direction. With this definition we compute the escape fraction for all
of our models, the results are shown in Fig \ref{figure:escape}\\

\begin{figure}
  \includegraphics[width=0.45\textwidth]{EscapeFraction.png}
    \label{figure:efvsv}\caption{Escape fraction for all the models. Left panels show the central distribution, while right panels show the homogeneous distribution} 
\end{figure}

In Figure \ref{figure:efvsv} we found that in the central distribution
the escape fraction does not change with velocity while it does in the
optical depth (See Verhamme 2006, an argument about this). On the other hand for the homogeneous distribution we
found that for higher velocities photons escape easily. The difference
between this two results relly in the fact that in the homogeneous
distribution photons are emitted closer to the escape surface and this
makes this configuration more sensitive to rotation while in the
central configuration the escape fraction depends mainly in the amount
of gas rather than in rotation.

As a final test we compare our results with the analytical solution of the slab developed by (Neufeld 1990) Fig. ~\ref{figure:efvsNeufeld}, as the geometry we use is different from the one described for Neufeld, we dont expect the same results, in fact we found that for the homogenous sphere the escape fraction is higer than the slab (see ref, mark). Also we notest that for $\tau<10^{6}$ the escape fraction does not increase as it will be expected, This is due to the fact that the condition $a\tau_{h}>>\tau_{d}$ in not valid anymore.

@@

@@

Untill now we have shown the main effects of gas and dust rotation in the Ly$\alpha$ line morphology such as the escape fraction, the FWHM the maxima position, also we see that shape of the outgoing spectra depends on the position of the observer. It is important to see if rotational effects are detectable in observational methods involving the Ly$\alpha$ line.

@@

\begin{table*}
\begin{center}
\begin{tabular}{cccc}\hline
Model & SSA22a 4980/80   & HS1549 4667/80 & HS1700 4018/90\\
\hline
V0HOM & 0.997 / 0.906 / 0.405 & 0.999 / 0.965 / 0.523 & 0.999 / 0.991 / 0.662 \\
V100HOM & 0.948 / 0.819 / 0.404 & 0.981 / 0.905 / 0.511 & 0.995 / 0.962 / 0.64 \\
V200HOM & 0.748 / 0.748 / 0.403 & 0.815 / 0.815 / 0.488 & 0.879 / 0.879 / 0.59 \\
V300HOM & 0.597 / 0.597 / 0.391 & 0.661 / 0.661 / 0.457 & 0.735 / 0.735 / 0.542 \\
V0HOMDust & 0.999 / 0.997 / 0.931 & 0.999 / 0.993 / 0.863 & 0.999 / 0.998 / 0.927\\
V100HOMDust & 0.994 / 0.985 / 0.912 & 0.986 / 0.967 / 0.863 & 0.996 / 0.989 / 0.9\\
V200HOMDust & 0.914 / 0.907 / 0.861 & 0.833 / 0.815 / 0.728 & 0.889 / 0.879 / 0.81\\
V300HOMDust & 0.831 / 0.826 / 0.799 & 0.676 / 0.668 / 0.618 & 0.749 / 0.744 / 0.702\\
\hline
\end{tabular}
\caption{
Values of fluxes for tree different NB and the our models simulated with CLARA
} 
\label{table:NBfilters}
\end{center}
\end{table*}

In table \ref{table:NBfilters} we present the results of the flux in every narrowband filter for the homogenous model at different velocities and hydrogen optical depth $\tau_{0}$. As is well known an increase in the optical depth makes the line peacks separation bigger. In fact for some cases the line width is larger than the NB filer width fig.(put ref fig) for those cases we modify the redshift in order to make the peak maxima match the NB filter center. For all cases we found that as velocity increase the flux is less. \\
In the case of the dusty model we found the same trend of the flux with the velocity but in this cases the effect is no that strong. 
We also found some dependency with the viewing angle.

@@

... The results derived in this paper have consequences on the
interpreation of galaxy observations in the Lyman alpha line.
..compare steidel et al (2011)
% taking into account equivalent widths results with narrow band
% filters we can make a criterium in order to select wich galaxie
% could be seen in actual observations.

\section{Conclusions}
In this paper we have estimated the effects of gas bulk rotation on
the emission  of the Lyman $\alpha$ line. We based the study on the
study of a simplified configuration of an homogeneous sphere rotating
as a solid body. We explored  a range of models by varying the
rotation speed, hydrogen optical depth, dust optical depth and initial
distribution of Lyman$\alpha$ photons with respect to the gas
density. This was implemented in CLARA, a Monte-Carlo
radiative transfer code already used to study the Lyman $\alpha$
line. \\

As first we see how the width of the line changes using a modified FWHM explained in section \ref{sec:widthpeak}, and we found that as gas bulk rotation incrase also the width increase in a factor of $2-3$ in comparison with the static case. we also take into account the influence of the observer position is and we found that for observers with a line of sight perpendicular to the axis of rotation measure a larger ($15\%$) line width than those aligned with the rotation axis.\\

As many observational Ly$\alpha$ (Kulas et al) spectra are double peaked  and those provide important information about the inflows/outflows of gas content, we study the effect of rotation in the position of this peaks and we find that the position of the maxima does not change with rotation.\\

One of the key features in the study of the Ly$\alpha$ line is the escape fraction, when rotational effects are taking into account we found that the escape fraction increase in about $20\%-30\%$ for the homogenous sphere model. (Nscatt)

4. Integrated flux.

>>>>

\begin{abstract}
Rotation is present in the gas kinematics of galaxies up to the
highest redshifts. In this paper we present for the first time
radiative transfer calculations that show the impact of rotation on
the morphology of the Lyman $\alpha$ line. To this end we construct
simplified models where a galaxy is modeled as an homogeneous sphere
composed as an homogeneous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with linear
velocities at the surface in the range $0-300$ \kms. We consider
radiation sources both in the center of the rotating cloud and also
homogeneously distributed around the sphere. We find that higher
rotational velocities increase the width of each peak in the outgoing
line profile while it also increases the amount of Lyman alpha photons
escaping in the line center. This trends makes that for high
rotational velocities and large Hydrogen optical depths the double
peak of the line tends to be erased an be replaced by a single peak the
lines center. This is more pronounced for radiation sources
homogeneously distributed. Concerning the escape fraction we find that
rotation does not have any effect, provided that all the sources are
centrally emitted. However in the case of homogeneously emitted
sources we measure an increase of about a factor of $2$ in the escape
fraction for higher rotational velocity values. 
Our work shows clearly that gas rotation has a non
negligible impact on the shape of the Lyman $\alpha$ line.

@@

Due to the resonant nature of the Lyman alpha line, gas kinematics
play an important role shaping its morphology. In the literature there
has been extensive studies of outflow/inflow configurations.

In this paper we study for the first time the impact of rotation on
the morphology of the Lyman $\alpha$ line. To isolate the effects of
rotation we focus on a simple system: the gas distribution is
spherical, with homogeneous density and the gas rotates as a solid
body. We base our work on the Monte Carlo base radiative transfer code
CLARA \citep{CLARA}.

@@

The central result of this paper is summarized in
Fig.~\ref{fig:differentvelocities} where we show that rotation has a
considerable effect on the morphology of the emergent Lyman-alpha line
both in the case where the photons are emitted at the sphere's center
and when they are initialized with an homogeneous distribution all
over the gas volume.

The results for these outgoing spectra are integrated over the whole
sphere, meaning that all the escaping photons were taking into account
regardless of the direction of the outgoing photons. Figure
\ref{fig:differentobservers} shows how if one gives a weight to each outgoing
photon according to its direction when escaping the gas distribution
it is possible to detect notable differences in the spectrum for
different viewing angles.

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{7tDifSpeedsZ.png}
  \includegraphics[width=0.45\textwidth]{7tHOMDifSpeeds1.png}
\caption{Shape of the lyman alpha line for
    different velocities. The left (right) panel shows the central
    (homogeneous) photon distribution. All photons were taken into
    account regardless of their outgoing direction of propagation.
  \label{fig:differentvelocities}}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.45\textwidth]{Observers7t.png}
\includegraphics[width=0.45\textwidth]{Observers7tHOM.png}
\caption{Same as Figure \ref{fig:differentvelocities} but this time
  the different lines show the line morphology for different polar angles
  $\theta$ with respect to the rotation axis. In this case each photon
  has a weight dependent on their outgoing direction. All spectra correspond
  to the same $V_{\rm max}=300$\kms, Optical   Depth $\tau=10^{7}$ and
  Central Distribution without dust.
  \label{fig:differentobservers}}

@@

In the following subsections we quantify the trends observed in
Fig. \ref{fig:differentvelocities} and
Fig. \ref{fig:differentobservers} as a function of the maximum
rotation velocity $V_{\rm max}$ and the position of the observer with
respect to the rotation angle, $\mu = \cos{\theta}$. All the results
in this section will be expressed in terms of rest frame wavelength.

@@

\subsection{Line width and peak maxima}
\label{sec:widthpeak}
\label{fig:differentvelocities}

\begin{figure}
    \includegraphics[width=0.45\textwidth]{WidthvsVmax.png}
\caption{Width of the lyman-alpha line for all the models.\label{fig:widthvsvelocity}}
\end{figure}

%\begin{figure*}%QUITARLA Y COMENTARLA
  %  \includegraphics[width=0.45\textwidth]{WidthvsThetaCentral.png}
   % \includegraphics[width=0.45\textwidth]{WidthvsTheta.png}
    %\caption{Width of the lyman-alpha line for all the models. 
 % \label{figure:widthvstheta}
%}
%\end{figure*}

The first quantitative conclusion of the effect of rotation in the
Lyman alpha line is that the double peaks in the line tend to broaden
until they reach a single broad emission peak. This is most evident in
the case of Lyman-alpha sources homogeneously distributed in the gas
distributions (Fig.~\ref{fig:differentvelocities} Right panel).

@@

%Im not sure about this anymore
This result includes all the outgoing photons, regardless of the
position of the observer. In Figure~\ref{fig:widthvstheta} we
take into account the different positions of the observer in the
measurement of the half-width $W_{1/2}$. From this we conclude
that observers with a line of sight perpendicular to the axis of
rotation (i.e. edge-on in the case of spiral galaxy) tend to measure
larger line widths than observers aligned with the rotation axis
(i.e. face-on). The influence of the observer position on the line
width, amounting always less than $15\%$ of a difference with respect
to the result that takes into account all the outgoing photons with
the same weight regardless of the relative observer position.

%Both for central and homogeneous distribution of sources, We found a
%dependency of the width with the rotation velocity of the gas cloud,
%this results are summarize in Figure ~\ref{figure:width}. As velocity
%increase also the equivalent width increase (for both homogeneous and
%central distribSi Limonution), it means that some photons escape with fewer
%scatterings than the static case but at the same time a few photons escape with% more scatterings, this can be seen clearly in
%Figure. (Nscatt)\\

%We also found a dependency with the viewing angles, in particular as
%the angle incre ase the width also increase Figure
%~\ref{fig:withvstheta}, this was also found in Verhemme et al. 2012
%for a rotating disk but the pure effect of rotation can not been study
%as it is in our case. Following the convention of Verhamme et al 2012
%we define $\mu=cos(\theta)$, and we can make a polynomial fit for the
%$FWHM(\mu)$ for our models, this is resumed in Table ~\ref{table:fits}

@@

In Figure~\ref{fig:maximumsvsvelocity} we present the position
of the maxima as a function of $V_{\rm max}$. For the photons emitted
in the line center we do not find any variation in the position of the
maxima in the range of explored parameter space (dash lines). However in the
homogeneous case we can see how the maxima goes to $x_{m}=0$, meaning
that the double peak is converted into a single peak.

This transition to a single peak line occurs for the systems where it
becomes easier for a bulk of the photons to escape with the lowest
number of scatterings possible. This can explain how the single peak
stage can be achieved in the homogeneous source distribution where
there is fraction of the photons inside a photosphere region with
$\tau_{H}\sim 1$ which allows them to escape within one
scatter. Increasing the rotational velocity makes it easier for the
photons in this photosphere region to escape.

We also take into account the viewing angle $\mu$ for all the optical 
depths and rotational velocities and the two different photon distribution.
But we don't find any dependency meaning that the peaks would be placed in
the same place no matter the position of an external observer.

%In Figure ~\ref{fig:maximavsangle} we present the position of the
%maxima as a function of the viewing angle $\mu$ for all the optical
%depths and rotational velocities and the two different photon
%distributions [FALTA LA GRAFICA].

@@

Finally we also report on the effect of the neutral Hydrogen
optical depth $\tau_{H}$ on the maxima position $x_{m}$
Figure~\ref{fig:maximumsvsvelocity}. We find that, at fixed
rotation velocity, the position of the maxima increases with optical
depth as expected from basic theoretical considerations. We compare
our results with the expected theoretical scaling for an infinite
slab.

@@

Until know we have focus in how rotation affect the morphology of the Lyman alpha line, now we turn to study deeply the possible causes of this effects. As recombination is the main cause of the emission of the Lyman alpha line, it is important to study the relationship between the number of times that a photon is absorbed and re emitted hereafter scatterings with it's shift in the wavelength.

As a first approach we study the number of scatterings of all the photons at different velocities and for both photon distributions Figure~\ref{fig:Nscatt}. For the central distribution we found that the average number of scattering does not have a relevant change ($\sim 0.5\%$) with $V_{max}$, For the Homogeneous distribution we found that as $V_{max}$ increase the average number of scatterings decrease in a$\sim 39\%$, it means that rotation makes photons to escape easier from the cloud. The reason why this is not observed in the central distribution is because the amount of gas that every photon has to break through is so large that the average number of scatterings is large too, making the effect of rotation unobservable.

In order to understand the effects of rotation in the morphology of the line we make histograms of $<N_{scatt}>$  in function of $X_{m}$ for both the central Fig.~\ref{fig:NscattHisto} and the  homogeneous distribution Fig.~\ref{fig:NscattHistoHOM}. For the central distribution and for lower velocities we found that for some value (put the value) of $N_{scatt}$ the mayority of photons have a defined value of $X_{m}$. While for higher velocities photons tend spread out in $X_{m}$(Make position analysis of this photons).\\

For the homogeneous distribution we found the same effect but also, for highest velocities the mayority of the photons near the surface escape with much less scatterings that in the static case. This analysis let us conclude that rotation make photons escape with less scatterings and it spread out the wavelenght of the of the outgoing photons represented by $X_{m}$.

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NscattvsVmax.png}
\caption{Log of the Average number of scatterings of the outgoing photons as function of the velocity. \label{fig:Nscatt}} 
\end{figure}

@@

Until now we have studied models in absence of dust, but is known (REFERENCE) that there is presence of dust in high redshift galaxies. Our models with dust are treated as is explained in detail in  the appendix of (Forero et al 2011).

Of particular interest is to compute the escape fraction of Lyman $\alpha$ photons coming from the most distant galaxies, due to the fact that with the observed intensity of the Lyman alpha line quantities as the LF and SFR can be derived (put some references).

Previous studies have shown the correlation of the Escape fraction with galactic mass (Laursen et al 2009, Dayal et al 2010) abundances and the kinematics of dust. In order to study pure rotational effects in the escape fraction we fixed the dust abundance $\tau_{A}=0.01$ and the galaxy mass. We compute the escape fraction for the models described in Table \ref{table:models}.

For a realistic model we also take into account the viewing angle, first we fixed the viewing angle at $\theta = 0$  Fig.~\ref{figure:efvsv} and then we fixed the velocity in order to see the escape fraction correlation with the viewing angle. Therefore we define the escape fraction as:\\

@@

Where NI is the initial number of photons and NF is the final,
$\vec{k}$ is the rotation axis direction and ${o}$ the observer
direction. With this definition we compute the escape fraction for all
of our models, the results are presented in Fig \ref{figure:efvsv}\\

\begin{figure}
  \includegraphics[width=0.45\textwidth]{EscapeFraction.png}
   \caption{Escape fraction for all the models. continious lines represent the homogenous models, while dashed lines represent the central models}\label{figure:efvsv}
\end{figure}

In Figure \ref{figure:efvsv} we found that in the central distribution
the escape fraction does not change with velocity while it does in the
optical depth (See Verhamme 2006, an argument about this). On the other
hand for the homogeneous distribution we
found that for higher velocities photons escape easily. The difference
between this two results rely in the fact that in the homogeneous
distribution photons are emitted closer to the escape surface and this
makes this configuration more sensitive to rotation while in the
central configuration the escape fraction depends mainly in the amount
of gas rather than in rotation.

As a final test we compare our results with the analytical solution of 
the slab developed by (Neufeld 1990) Fig. ~\ref{figure:efvsNeufeld}, as 
the geometry we use is different from the one described for Neufeld, we 
dont expect the same results, in fact we found that for the homogenous 
sphere the escape fraction is higer than the slab (see ref, mark). Also 
we notest that for $\tau<10^{6}$ the escape fraction does not increase 
as it will be expected, This is due to the fact that the condition 
$a\tau_{h}>>\tau_{d}$ in not valid any more.

@@

\subsection{Inhomogeneous model}

Now we set up a test model when the photons are placed in an inhomogeneous model,

@@

Until now we have shown the main effects of gas and dust rotation in the Ly$\alpha$ line morphology such as the escape fraction, the FWHM, the maxima position, also we see that shape of the outgoing spectra depends on the position of the observer. It is important to see if rotational effects are detectable in observational methods involving the Ly$\alpha$ line.

@@

\begin{table*}
\begin{center}
\begin{tabular}{cccc}\hline
Model & SSA22a 4980/80   & HS1549 4667/80 & HS1700 4018/90\\
\hline
V0HOM & 0.99 / 0.90 / 0.40 & 1.0 / 0.96 / 0.52 & 1.0 / 0.99 / 0.66 \\
V100HOM & 0.94 / 0.81 / 0.40 & 0.98 / 0.90 / 0.51 & 0.99 / 0.96 / 0.64 \\
V200HOM & 0.74 / 0.74 / 0.40 & 0.81 / 0.81 / 0.48 & 0.87 / 0.87 / 0.59 \\
V300HOM & 0.59 / 0.59 / 0.39 & 0.66 / 0.66 / 0.45 & 0.73 / 0.73 / 0.54 \\
V0HOMDust & 0.99 / 0.99 / 0.93 & 0.99 / 0.99 / 0.86 & 0.99 / 0.99 / 0.92\\
V100HOMDust & 0.99 / 0.98 / 0.91 & 0.98 / 0.96 / 0.86 & 0.99 / 0.98 / 0.9\\
V200HOMDust & 0.91 / 0.90 / 0.86 & 0.83 / 0.81 / 0.72 & 0.88 / 0.87 / 0.81\\
V300HOMDust & 0.83 / 0.82 / 0.79 & 0.67 / 0.66 / 0.61 & 0.74 / 0.74 / 0.70\\
\hline
\end{tabular}
\caption{
Values of fluxes for tree different NB and the our models simulated with CLARA
} 
\label{table:NBfilters}
\end{center}
\end{table*}

In table \ref{table:NBfilters} we present the results of the flux in every 
narrowband filter for the homogeneous model at different velocities and hydrogen 
optical depth $\tau_{0}$. As is well known an increase in the optical depth makes 
the line peaks separation bigger. In fact for some cases the line width is larger 
than the NB filer width fig.(put ref fig) for those cases we modify the redshift 
in the available range in order to make the peak maxima match the NB filter center. 
For all cases we found that as velocity increase the flux is less.

In the case of the dusty model we found the same trend of the flux with the velocity 
but in this cases the effect is no that strong. 
We also found some dependency with the viewing angle.

\begin{figure}
  \includegraphics[width=0.45\textwidth]{NB7tDifVHOM.png}
 \label{figure:efvsNeufeld}\caption{NB filter located at the center of the maxima, with different redshift} 
\end{figure}

@@

... The results derived in this paper have consequences on the
interpretation of galaxy observations in the Lyman alpha line.
..compare steidel et al (2011)
% taking into account equivalent widths results with narrow band
% filters we can make a criterium in order to select wich galaxie
% could be seen in actual observations.

\section{Conclusions}
In this paper we have estimated the effects of gas bulk rotation on
the emission  of the Lyman $\alpha$ line. We based the study on the
study of a simplified configuration of an homogeneous sphere rotating
as a solid body. We explored  a range of models by varying the
rotation speed, hydrogen optical depth, dust optical depth and initial
distribution of Ly$\alpha$ photons with respect to the gas
density. This was implemented in CLARA, a Monte-Carlo
radiative transfer code already used to study the Lyman $\alpha$
line.

As first we see how the width of the line changes using a modified FWHM
explained in section \ref{sec:widthpeak}, and we found that as gas bulk 
rotation increase also the width increase in a factor of $2-3$ in comparison 
with the static case. We also take into account the influence of the observer 
viewing angle, we found that observers with a line of sight perpendicular  
to the axis of rotation measure a $15\%$ larger line width than those 
aligned with the rotation axis.

As many observational spectra Ly$\alpha$ emission line (Kulas et al) is double 
peaked, these peaks provide important information concerning gas kinematics and geometry, 
which can be partially explained with inflows/outflows of gas content. 
We study the effect of rotation in the position of this peaks, and we find 
that the position of the maxima does change with rotation for the homogeneous 
models when the double peak merged into a single peak as velocity increase. 
This effect is not seen for the central distribution when the double peak 
remains constant as the velocity increase. We also find that there is no 
dependency in the observer viewing angle with the maxima position.

Concerning the escape fraction under rotational effects on the Ly$\alpha$ 
emission line, we found that the escape fraction increase in about $20\%-30\%$
for the homogeneous sphere model. While rotational effects are negligible for
the central models and the escape fraction remains constant. Also the observer
viewing angle have no effect in the escape fraction neither for the homogeneous
and central models. Complementing this analysis we study the average number of 
scatterings $<N_{scatt}>$ that photons perform before escaping of the cloud
taking into account rotational effects. The main result here is for the homogeneous
models for which as velocity increase photons escape with about $\sim 39\%$ 
less scatterings than in the static case.

As an application of these results we compute the integrated flux taking 
into account the narrow band filters used by (Steidel et al 2011), for 
our models we found an important decrease up to $~40\%$ for the homogeneous models, 
and up to $22\%$ for the dusty homogeneous models in the flux as velocity increase. 
Also we calculate at what redshift should the filter be in order to get the maximum 
flux, and for the tree filters we get values that rely in the filter redshift range.
This effects would have a relevant implication at the time to find high redshift 
galaxies.

This paper illustrates for the first time the main effects of rotation in  
the morphology of the Ly$\alpha$ emission line, we estimate the range of this effects
for simplified models.
